---
title: "Coding_Challenge4_Markdown"
author: "Pankaj Gaonkar"
date: "2025-02-27"
output:
  pdf_document:
  md_document:
    variant: gfm
  word_document:
  html_document:
    toc: true
    toc_float: true
---

# Coding Challenge 4 Answers

Q1) Explain the following
a.	YAML header
YAMLS helps in writing the configuration of the files and is on the top of the markdown files.


b.	Literate programming
In literate programming along with our code we will have its explanations and provide the chunks of codes. Itexplains the codes in bettwe way.

Q2)

## Below is the clickable link to manuscript where these data are published

[Noel et al., 2022. Endophytic fungi as promising biocontrol agent to protect wheat from Fusarium graminearum head blight. Plant Disease](https://doi.org/10.1094/PDIS-06-21-1253-RE)


```{r coding}
## Load libraries
library(tidyverse)
library(ggpubr)
library("knitr")  # required for knitting


## Reading data by reative path

csv <- read.csv("MycotoxinData.csv", na.strings = "na")
head (csv)

## Codes needed for complete analysis

#Plot1 Treatment on Y

P1 <- ggplot(csv, aes(x = Treatment, y = DON, fill = Cultivar))+
  theme_classic() +                # removes the grids
  geom_boxplot() + 
  xlab("") +
  ylab("DON (ppm)") +
  geom_point(alpha = 0.6, shape = 21, position = position_jitterdodge()) +
  scale_fill_manual (values = c("#56B4E9", "#009E73")) +
  facet_wrap(~ Cultivar) 

str(csv)

csv$Treatment <- factor(csv$Treatment, levels = c("NTC", "Fg","Fg + 37", "Fg + 40", "Fg + 70"))


#Plot2 X15ADON on Y

P2 <- ggplot(csv, aes(x = Treatment, y = X15ADON, fill = Cultivar))+
  theme_classic() +                # removes the grids
  geom_boxplot() + 
  xlab("") +
  ylab("15ADON") +
  geom_point(alpha = 0.6, shape = 21, position = position_jitterdodge()) +
  scale_fill_manual (values = c("#56B4E9", "#009E73")) +
  facet_wrap(~ Cultivar) 


#Plot3 MassperSeed_mg on Y

P3 <- ggplot(csv, aes(x = Treatment, y = MassperSeed_mg, fill = Cultivar))+
  theme_classic() +                # removes the grids
  geom_boxplot() + 
  xlab("") +
  ylab("Seed Mass (mg)") +
  geom_point(alpha = 0.6, shape = 21, position = position_jitterdodge()) +
  scale_fill_manual (values = c("#56B4E9", "#009E73")) +
  facet_wrap(~ Cultivar)
```


## Code from coding challenge 3, question 5 & sepreated code chunks for each plit


```{r P1}

P1_stat <- P1 + 
            geom_pwc(aes(group = Treatment), method = "t_test", label = "p.adj.format")
P1_stat

```

```{r P2}
P2_stat <- P2 + 
  geom_pwc(aes(group = Treatment), method = "t_test", label = "p.adj.format")
P2_stat
```

```{r P3}
P3_stat <- P3 + 
  geom_pwc(aes(group = Treatment), method = "t_test", label = "p.adj.format")
P3_stat
```



```{r Combined}
#combining
Combined_plot_stat <- ggarrange(
  P1_stat,   ## add the plots onf interest subsequently
  P2_stat,
  P3_stat,
  labels = "auto",  # Automatically label the plots (A, B, C, etc.)
  nrow = 1,  # Arrange the plots in # rows
  ncol = 3,  # Arrange the plots in # column
  common.legend = T
)


```

### GitHub link:
https://github.com/ppg0001/PLPA_Assignment/tree/main

