---
title: "Coding_Practice_Linear_Model"
author: "Pankaj Gaonkar"
date: "2025-04-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Linear Modelling

```{r}
#Load packages
library(tidyverse)
library(lme4)

#install.packages("multcompView")
library(multcomp)

#install.packages("emmeans")
library(emmeans)

```


```{r}
data("mtcars")

#The first step is always to visualize your data.

#If you have two continuous variables, you will want a scatter plot
ggplot(mtcars, aes(x = wt, y = mpg)) +    #plotting weight against mpg
  geom_smooth(method = lm, se = FALSE, color = "grey") +
  geom_point(aes(color = wt)) +
  xlab("Weight") + 
  ylab("Miles per gallon") +
  scale_colour_gradient(low = "forestgreen", high = "black") +
  theme_classic()

```
#Does it look like there is a relationship? But how do we know?
#By running the liear model (lm)

```{r}

#Continious Y and continious X example. Runnig linear model we would call it a regression

lm(mpg ~ wt, data = mtcars)

lm1 <- lm(mpg ~ wt, data = mtcars)

summary(lm1)

```


```{r}
#Correlation analysis- it will be it a cause-and-effect relationship.

cor.test(mtcars$wt, mtcars$mpg)


```


```{r}
#Getting residuals from linear model
model <- lm(mpg~wt, data = mtcars)

ggplot(model, aes(y = .resid, x = .fitted)) +
  geom_point() +
  geom_hline(yintercept = 0)


```
#We will use bull_richness.csv data for further analysis
#Note: poisonn better for count data (example richness variable has all integer, its a count data. In ANOVA assumed distribution is gausiaan distribution that can take any value between negative infinity and infinity but integer values dont fit that, so poisson better for our variable of richness. BUT WE WILL JUST GO AHEAD WITH REGULAR THINGY AND WONT DO glm-poissson)

```{r}
#Load data

bull.rich <- read.csv("Bull_richness.csv")

```


```{r}
bull.rich.sub <- bull.rich %>%
  filter(GrowthStage == "V8" & Treatment == "Conv.")

t.test(richness~Fungicide, data = bull.rich.sub)  #For just running t-test

summary(lm(richness~Fungicide, data = bull.rich.sub))  #runnig as linear model
  

```

## ANOVA

## ANOVAs would have continuous y and multinomial categorical x. Meaning, more than two groups.


```{r}
#Filtering dataset

bull.rich.sub2 <- bull.rich %>%
  filter(Fungicide == "C" & Treatment == "Conv." & Crop == "Corn")

bull.rich.sub2$GrowthStage <- factor(bull.rich.sub2$GrowthStage, levels = c("V6", "V8", "V15"))

ggplot(bull.rich.sub2, aes(x = GrowthStage, y = richness)) +
  geom_boxplot()


```


```{r}
#Now lets set up LINEAR MODELS instead of ANOVA. 
#Same result as ANOVA because ANOVA is a regression.

lm.growth <- lm(richness ~ GrowthStage, data = bull.rich.sub2)
summary(lm.growth)

anova(lm.growth)
summary(aov(richness ~ GrowthStage, data = bull.rich.sub2))


```

## how do we know which is different from the the other group? By post-hoc. Using 'means' and 'multcomp' package

```{r}
#Using MEANS

lsmeans <- emmeans(lm.growth, ~GrowthStage) # estimate lsmeans
Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE) # contrast with Tukey ajustment by default. 
Results_lsmeans



```
## Interaction terms

```{r}
#Filter dataset
bull.rich.sub3 <- bull.rich %>%
  filter(Treatment == "Conv." & Crop == "Corn")

bull.rich.sub3$GrowthStage <- factor(bull.rich.sub3$GrowthStage, levels = c("V6", "V8", "V15"))

```


```{r}
# Set up our linear model with fungicide interaction factor
#Style 1
lm.inter <- lm(richness ~ GrowthStage + Fungicide + GrowthStage:Fungicide, data = bull.rich.sub3)
summary(lm.inter)

#Style 2
lm(richness ~ GrowthStage*Fungicide, data = bull.rich.sub3)
summary(lm(richness ~ GrowthStage*Fungicide, data = bull.rich.sub3))

anova(lm.inter) # The interaction term is significant. But why to run this again?



```


```{r}
# For this study we more care about the effect of fungicide over the levels of growthstage so we can plot that easily.
bull.rich.sub3 %>%
  ggplot(aes(x = GrowthStage, y = richness, fill = Fungicide)) +
  geom_boxplot()

#Note: So, by visualization, we can see that it looks like the V8 fungicide affected richness more than others, and by V15, they recovered.But how do we know? In this case, we have to do a post-hoc test within the levels of growthstage.

```

# Mixed effects models
#Important things to understand is 'fixed and random effects terms'. 

```{r}
lme0 <- lm(richness ~ GrowthStage*Fungicide, data = bull.rich.sub3)

#How to include random effect in 'lme0'?
#using lmer function

lme1 <- lmer(richness ~ GrowthStage*Fungicide + (1|Rep), data = bull.rich.sub3)
summary(lme1)


```


```{r}
summary(lme0)

summary(lme1)
```


```{r}

```


















