---
title: "Iteration, Loops & Function_Practice"
author: "Pankaj Gaonkar"
date: "2025-03-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#load packages

#install.packages("drc")

library(drc)
library(tidyverse)
library(ggplot2)


```

# Fuctions
## Functions are like those written into packages. They are useful when you need to perform the same code on different data and you want to avoid copy and paste errors.

```{r}
# Function to convert Fahrenheit to Celsius. Lets say degree_f=Fahrenheit
#Example (5*(degree_f - 32)/9)   < formula for Fahrenheit to Celsius

(5*(32 - 32)/9) #if degree_f is 32

(5*(80 - 32)/9) # if degree_f is 80

(5*(40 - 32)/9) # if degree_f is 40

#But then we end up doing lots of copy paste everytime we want to convert the temperature, prone to errors. So to avoid this we can use the "Functions".

```

# Converting it to a function
```{r}
# I broke it down in step by step. Just for my understanding

#Step 1
F_to_C <- function(f_temp){      
  (5*(f_temp - 32)/9)       #What will happen is when we input a value for f_temp for "F_to_C" we will get the answer to formula in function.
}

#Step 2 (This is complete function)
F_to_C <- function(f_temp){      
  celsius <- (5*(f_temp - 32)/9)    # we are saving value in the celsius
  return(celsius)
}

#Step3_using/calling function

F_to_C(32) #Here we wil get the 32F to its celsius value.
F_to_C(82)


```

# Task _ Write a function to convert celsius to fahrenheiht using the conversion of F = C x (9/5)+32
```{r}

C_to_F <- function(c_temp){
  farhenheit <- (c_temp*(9/5)+32)
  return(farhenheit)
}

C_to_F(0)

```

#Iterations
1. rep() 

2. seq() & seq_along() 

3. for loops map() *nested map()

```{r}
#Repetition
#rep() COMMAND

rep("A", 3) # repeats A three times 

rep(c("A", "B"), 5) # repeats A and B, times

rep(c(1,2,3,4), times = 4) # repeats 1,2,3,4, 4 times 

rep(c(1,2,5,2), times = 4, each = 4) # repeats 1 four times, 2 four times, 5 four times, and 2 four times. 


```
```{r}
#sequences of numbers easily
# seq() COMMAND

seq(from = 1, to = 7) # sequence of numbers 1 to 7

seq(from = 0, to = 10, by = 2) # sequence of numbers from 0 to 10 by 2s 

rep(seq(from = 0, to = 10, by = 2), times = 3, each = 2) # combined seq() and rep()


```


```{r}
# The seq_along() function 
## It allows you to generate a sequence of numbers based on non-integer values. This will become very useful when we want to loop over elements within a dataframe.

LETTERS  # # use the built in LETTERS vector for an example. 

seq_along(LETTERS)
seq_along(LETTERS[1:5]) # will return 1,2,3,4,5 not the actual letters. 
#This will help in loops, knwoing the sequence number in the vector while looping. To be continued in loop section.


```

# The "For" loop

```{r}
for (i in 1:10){    # i commonly used can be anything, we will define our iteration
  print(i*2)        #for each value of i will be iterated till it gets to 10 while multiplying with 2.
}


```



Complicated example combining function in our loop.

```{r}

# Note: F_to_C is the function that we already defined earlier for temperature conversiont

for(i in -30:100){     #setting up sequence
  result <- F_to_C(i)   #For each value in sequence will substituted in i as iteration
  print(result)
}

```

This was great, but this only printed to the console, and we cannot really do anything with those values. What if we wanted to do something with the result of the iteration?

Step 1. Set a R object to NULL

Step 2. Set your for loop

Step 3. Save the result of your for loop into a dataframe each iteration

Step 4. append one row of the dataframe to the null object each iteration of the loop. 


```{r}

celcius.df <- NULL 
for (i in -30:100){   
  result <- data.frame(F_to_C(i), i)   #create one row dataframe called data.frame contain two columns "F_to_C(i)" and "i".
  celcius.df <- rbind.data.frame(celcius.df, result)  #binding celsius.df to result

}


celcius.df

```















