---
title: "Coding_Challenge_DataWrangling"
author: "Pankaj Gaonkar"
date: "2025-03-20"
output: 
  pdf_document:
  md_document:
    variant: gfm
  word_document:
  html_document:
    toc: true
    toc_float: true
---

# Coding Challenge _ Data Wrangling


```{r Packages}
#install.packages("tidyverse")
library(tidyverse)

```


# 1.	3 pts. Download two .csv files from Canvas called DiversityData.csv and Metadata.csv, and read them into R using relative file paths.

```{r load data}
DiversityData <- read.csv("DiversityData.csv")
str(DiversityData)
```


```{r}
Metadata <- read.csv("Metadata.csv")

str(Metadata)
```





# 2.	4 pts. Join the two dataframes together by the common column ‘Code’. Name the resulting dataframe alpha.

```{r}

alpha  <- left_join(Metadata, DiversityData, by = "Code")

```


# 3.	4 pts. Calculate Pielou’s evenness index: Pielou’s evenness is an ecological parameter calculated by the Shannon diversity index (column Shannon) divided by the log of the richness column. 
#a.	Using mutate, create a new column to calculate Pielou’s evenness index. 
#b.	Name the resulting dataframe alpha_even.

```{r}

# Creating a new column
alpha_even  <-  mutate(alpha, Pielou_eveness = shannon / log(richness))


```



# 4 4.	Pts. Using tidyverse language of functions and the pipe, use the summarise function and tell me the mean and standard error evenness grouped by crop over time.
#a.	Start with the alpha_even dataframe
#b.	Group the data: group the data by Crop and Time_Point.
#c.	Summarize the data: Calculate the mean, count, standard deviation, and standard error for the even variable within each group.
#d.	Name the resulting dataframe alpha_average



```{r}

#summary statistics

alpha_average  <-  alpha_even %>%
    group_by(Crop, Time_Point) %>%
    summarise(mean.even = mean(Pielou_eveness), # calculating the mean richness, stdeviation, and standard error
            n = n(), 
            sd.dev = sd(Pielou_eveness)) %>%
    mutate(std.err = sd.dev/sqrt(n))



```


# 5.	4. Pts. Calculate the difference between the soybean column, the soil column, and the difference between the cotton column and the soil column
#a.	Start with the alpha_average dataframe
#b.	Select relevant columns: select the columns Time_Point, Crop, and mean.even.
#c.	Reshape the data: Use the pivot_wider function to transform the data from long to wide format, creating new columns for each Crop with values from mean.even.
#d.	Calculate differences: Create new columns named diff.cotton.even and diff.soybean.even by calculating the difference between Soil and Cotton, and Soil and Soybean, respectively.
#e.	Name the resulting dataframe alpha_average2



```{r}

alpha_average2 <- alpha_average %>%
              select(Time_Point, Crop, mean.even) %>%
              pivot_wider(names_from = Crop, values_from = mean.even) %>%
              mutate(diff.cotton.even = Soil - Cotton) %>%
              mutate(diff.soybean.even = Soil - Soybean)
  
  
  
```



# 6.	4 pts. Connecting it to plots
#a.	Start with the alpha_average2 dataframe
#b.	Select relevant columns: select the columns Time_Point, diff.cotton.even, and diff.soybean.even.
#c.	Reshape the data: Use the pivot_longer function to transform the data from wide to long format, creating a new column named diff that contains the values from diff.cotton.even and diff.soybean.even.
#i.	This might be challenging, so I’ll give you a break. The code is below.

#pivot_longer(c(diff.cotton.even, diff.soybean.even), names_to = "diff")

#d.	Create the plot: Use ggplot and geom_line() with ‘Time_Point’ on the x-axis, the column ‘values’ on the y-axis, and different colors for each ‘diff’ category. The column named ‘values’ come from the pivot_longer. The resulting plot should look like the one to the right. 


```{r}

alpha_average2 %>%
  select(Time_Point, diff.cotton.even, diff.soybean.even) %>%
  pivot_longer(c(diff.cotton.even, diff.soybean.even), names_to = "diff") %>%
  ggplot(aes(x = Time_Point, y = value, color=diff)) + # Plot it 
  geom_line() +
  theme_minimal() +
  xlab("Time (hrs)") +
  ylab("Difference from soil in Pielou’s evenness")

```



## Below is the clickable link to GitHub Folder where these files are published

[Coding_challenge_5 Folder](https://github.com/ppg0001/PLPA_Assignment/tree/main/Coding_challenge_5)


