---
title: "Coding_challenge_7"
author: "Pankaj Gaonkar"
date: "2025-04-03"
output:
  pdf_document:
  md_document:
    variant: gfm
  word_document:
  html_document:
    toc: true
    toc_float: true
---

1.	4 pts. Read in the data called “PlantEmergence.csv” using a relative file path and load the following libraries. tidyverse, lme4, emmeans, multcomp, and multcompView. Turn the Treatment , DaysAfterPlanting  and Rep into factors using the function as.factor
STAND$Treatment <- as.factor(STAND$Treatment) # example shown here. 

```{r}
#load libraries
#Load packages
library(tidyverse)
library(lme4)

#install.packages("multcompView")
library(multcomp)

#install.packages("emmeans")
library(emmeans)
```



```{r}
#load data
datum <- read.csv("PlantEmergence.csv")
head(datum)

```

```{r}
#Turn the Treatment , DaysAfterPlanting  and Rep into factors using the function as.factor

datum$Treatment <- as.factor(datum$Treatment)
datum$DaysAfterPlanting <- as.factor(datum$DaysAfterPlanting)
datum$Rep <- as.factor(datum$Rep)


```



2.	5 pts. Fit a linear model to predict Emergence using Treatment and DaysAfterPlanting along with the interaction. Provide the summary of the linear model and ANOVA results. 

```{r}
model1 <- lm(Emergence ~ Treatment + DaysAfterPlanting + Treatment:DaysAfterPlanting, data = datum)
summary(model1)

anova(model1)


```


3.	5 pts. Based on the results of the linear model in question 2, do you need to fit the interaction term? Provide a simplified linear model without the interaction term but still testing both main effects. Provide the summary and ANOVA results. Then, interpret the intercept and the coefficient for Treatment 2.

## Answer:
**None of the interactions were significant so we can exclude the interactions. Lets consider with-interaction as 'complicated model'(model 1) and without-interaction as 'simple model'(model 2). So we will proceed with simple model for our analysis.**

**Interpretation:**

- **Intercept is showing that Emergence value is 182.163 units when other independet variables(Treament, DaysAfterPlanting) are considered to be zero.**

- **Treatment 2: Plant that received treatment 2 had 134.531 units (SE +/- 3.42) lesser emergence. (p < < 2e-16)**



```{r}

model2 <- lm(Emergence ~ Treatment + DaysAfterPlanting, data = datum)
summary(model2)
anova(model2)

```


4.	5 pts. Calculate the least square means for Treatment using the emmeans package and perform a Tukey separation with the compact letter display using the cld function. Interpret the results.


## Answer:

** Interpretation **
**Based on multiple comparison we found that statistically significant differences were only observed between Treatment 2 and other treatment groups.There was lower Emergence in group 2 compared to other groups.**


```{r}

lsmeans <- emmeans(model2, ~ Treatment) # estimate lsmeans
Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE) 
Results_lsmeans

```




5.	4 pts. The provided function lets you dynamically add a linear model plus one factor from that model and plots a bar chart with letters denoting treatment differences. Use this model to generate the plot shown below. Explain the significance of the letters. 


## Answer:
**Significance of letters: if the letters are same then there is no statically significant difference between groups. If letters are different between two groups then they are statistically signifcantly different from each other.

**Example:**
**Treament 2 (letter-c) versus Treamet 9 (letter- ab) -- They are statically significantly different from each other.**

**Treatment 8 (letter-ab) versus Treament 9 (letter- ab) - There is NO statiically signficant differnce between them**

# Function defined (from assignment)

```{r}
plot_cldbars_onefactor <- function(lm_model, factor) {
  data <- lm_model$model
  variables <- colnames(lm_model$model)
  dependent_var <- variables[1]
  independent_var <- variables[2:length(variables)]

  lsmeans <- emmeans(lm_model, as.formula(paste("~", factor))) # estimate lsmeans 
  Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE, Letters = letters) # contrast with Tukey adjustment by default.
  
  # Extracting the letters for the bars
  sig.diff.letters <- data.frame(Results_lsmeans$emmeans[,1], 
                                 str_trim(Results_lsmeans$emmeans[,7]))
  colnames(sig.diff.letters) <- c(factor, "Letters")
  
  # for plotting with letters from significance test
  ave_stand2 <- lm_model$model %>%
    group_by(!!sym(factor)) %>%
    dplyr::summarize(
      ave.emerge = mean(.data[[dependent_var]], na.rm = TRUE),
      se = sd(.data[[dependent_var]]) / sqrt(n())
    ) %>%
    left_join(sig.diff.letters, by = factor) %>%
    mutate(letter_position = ave.emerge + 10 * se)
  
  plot <- ggplot(data, aes(x = !! sym(factor), y = !! sym(dependent_var))) + 
    stat_summary(fun = mean, geom = "bar") +
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
    ylab("Number of emerged plants") + 
    geom_jitter(width = 0.02, alpha = 0.5) +
    geom_text(data = ave_stand2, aes(label = Letters, y = letter_position), size = 5) +
    xlab(as.character(factor)) +
    theme_classic()
  
  return(plot)
}

```




```{r}

plot_cldbars_onefactor(model2, "Treatment") #I realized functions can make life easier

```




6.	2 pts. Generate the gfm .md file along with a .html, .docx, or .pdf. Commit, and push the .md file to github and turn in the .html, .docx, or .pdf to Canvas. Provide me a link here to your github.

[Coding_challenge_7 Folder](https://github.com/ppg0001/PLPA_Assignment/tree/main/Coding_challenge_7)



