---
title: "CodingChallenge6_Loops and iterations"
author: "Pankaj Gaonkar"
date: "2025-03-27"
output: 
  pdf_document:
  md_document:
    variant: gfm
  word_document:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. 2 pts. Regarding reproducibility, what is the main point of writing your own functions
and iterations?

Answer: They are useful when you need to perform the same code on different data and you want to avoid copy and paste errors. Especially with function we can specify the formula to be performed and with iterations we can have several repetitions which wont be practically possible to write without error. Overall they help in reproducibility by avoiding the errors.



2. 2 pts. In your own words, describe how to write a function and a for loop in R and how
they work. Give me specifics like syntax, where to write code, and how the results are
returned

Answer:
Example:

We write a code in R chunk in R markdown.
Example:

## Function:

```

F_to_C <- function(f_temp){      
  celsius <- (5*(f_temp - 32)/9)    # we are saving value in the celsius
  return(celsius)                   # gives the answer to the input after running the functions. We get Farheneit converted to censius.
}

```

*Function is defined as F_to_C
*f_temp is our input

*Calling the function by giving input as 32 for this example:
*F_to_C(32) #Here we wil get the 32F to its celsius value.


## Loop:

Example of loop:

```
celcius.df <- NULL 
for (i in -30:100){   
  result <- data.frame(F_to_C(i), i)   #create one row dataframe called data.frame contain two columns "F_to_C(i)" and "i".
  celcius.df <- rbind.data.frame(celcius.df, result)  #binding celsius.df to result

}

celcius.df
```

*We start by creating a NULL object so that we can call that dataframe and we can have to output values.
*for (i in -30:100) -- This is a for loop where it will go through every value between -30 tp 100 Farheneit.
*We create results as data.frame as per desired column arrangement.
*Finally we input the results in the initial NULL dataframe.
*cecius,df will print out our entire dataframe with our loop output.




```{r}
#load packages

library("dplyr")

```





3. 2 pts. Read in the Cities.csv file from Canvas using a relative file path.

```{r}
#load data

datum <- read.csv("Cities.csv")

head(datum)


```


4. 6 pts. Write a function to calculate the distance between two pairs of coordinates based
on the Haversine formula (see below). The input into the function should be lat1, lon1,
lat2, and lon2. The function should return the object distance_km. All the code below
needs to go into the function.



```{r}

Distance_pair <- function(lat1, lon1, lat2, lon2){      
  # convert to radians
  rad.lat1 <- lat1 * pi/180
  rad.lon1 <- lon1 * pi/180
  rad.lat2 <- lat2 * pi/180
  rad.lon2 <- lon2 * pi/180
  # Haversine formula
  delta_lat <- rad.lat2 - rad.lat1
  delta_lon <- rad.lon2 - rad.lon1
  a <- sin(delta_lat / 2)^2 + cos(rad.lat1) * cos(rad.lat2) * sin(delta_lon / 2)^2
  c <- 2 * asin(sqrt(a))
  # Earth's radius in kilometers
  earth_radius <- 6378137
  # Calculate the distance
  distance_km <- (earth_radius * c)/1000

  return(distance_km)
}

```



5. 5 pts. Using your function, compute the distance between Auburn, AL and New York
City
a. Subset/filter the Cities.csv data to include only the latitude and longitude values
you need and input as input to your function.
b. The output of your function should be 1367.854 km



```{r}
# 5a) Subsetting data to include only new york and auburn

  lat1 <- datum$lat[datum$city_ascii == "Auburn"]
  lon1 <- datum$long[datum$city_ascii == "Auburn"]
  
  lat2 <- datum$lat[datum$city_ascii == "New York"]
  lon2 <- datum$long[datum$city_ascii == "New York"]

Distance_pair(lat1, lon1, lat2, lon2)



```

6. 6 pts. Now, use your function within a for loop to calculate the distance between all other
cities in the data

```{r}
#looping


Distance_all_cities.df <- NULL 

for (i in datum$city_ascii){
  lat2 <- datum$lat[datum$city_ascii == i]
  lon2 <- datum$long[datum$city_ascii == i]
  
  result <- Distance_pair(lat1, lon1, lat2, lon2)
  
  Distance_all_cities.df <- rbind.data.frame(Distance_all_cities.df, result)
  

}

Distance_all_cities.df


```


Q) Bonus point if you can have the output of each iteration append a new row to a
dataframe, generating a new column of data. In other words, the loop should create a
dataframe with three columns called city1, city2, and distance_km, as shown below. The
first six rows of the dataframe are shown below

```{r}
#Bonus

Distance_all_cities.df <- NULL 

for (i in datum$city_ascii){
  lat2 <- datum$lat[datum$city_ascii == i]
  lon2 <- datum$long[datum$city_ascii == i]
  
  result <- Distance_pair(lat1, lon1, lat2, lon2)
  
  Combined <- data.frame(city1 = "Auburn", City2 = i, distance_km = result)   #important to specify the dataframe name
  
  Distance_all_cities.df <- rbind.data.frame(Distance_all_cities.df, Combined)
  
}

Distance_all_cities.df

```



7. 2 pts. Commit and push a gfm .md file to GitHub inside a directory called Coding Challenge 6.
Provide me a link to your github written as a clickable link in your .pdf or .docx










