---
title: "CodingChallenge6_Loops and iterations"
author: "Pankaj Gaonkar"
date: "2025-03-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




1. 2 pts. Regarding reproducibility, what is the main point of writing your own functions
and iterations?




2. 2 pts. In your own words, describe how to write a function and a for loop in R and how
they work. Give me specifics like syntax, where to write code, and how the results are
returned







3. 2 pts. Read in the Cities.csv file from Canvas using a relative file path.

```{r}
#load data

datum <- read.csv("Cities.csv")

head(datum)


```


4. 6 pts. Write a function to calculate the distance between two pairs of coordinates based
on the Haversine formula (see below). The input into the function should be lat1, lon1,
lat2, and lon2. The function should return the object distance_km. All the code below
needs to go into the function.



```{r}

Distance_pair <- function(lat1, lon1, lat2, lon2){      
  # convert to radians
  rad.lat1 <- lat1 * pi/180
  rad.lon1 <- lon1 * pi/180
  rad.lat2 <- lat2 * pi/180
  rad.lon2 <- lon2 * pi/180
  # Haversine formula
  delta_lat <- rad.lat2 - rad.lat1
  delta_lon <- rad.lon2 - rad.lon1
  a <- sin(delta_lat / 2)^2 + cos(rad.lat1) * cos(rad.lat2) * sin(delta_lon / 2)^2
  c <- 2 * asin(sqrt(a))
  # Earth's radius in kilometers
  earth_radius <- 6378137
  # Calculate the distance
  distance_km <- (earth_radius * c)/1000

  return(distance_km)
}

```


```{r}
```





